---
title: "Day 20 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readr)
library(collections)
library(reshape2)
options(scipen = 999)
```

```{r,eval=FALSE}
input<-read_lines("Day20Sample.txt")

```

Put the particles into a dataframe
```{r,eval=FALSE}
particleframe<-as.data.frame(matrix(nrow=0,ncol=10))
for(i in 1:length(input)){
  particle<-input[i]
  particle<-str_remove_all(particle,"p=<|a=<|v=<|>,|>")
  particle<-c(i-1,as.numeric(unlist(str_split(particle,",| "))))
particleframe<-rbind(particleframe,particle)
}
colnames(particleframe)<-c("num","px","py","pz","vx","vy","vz","ax","ay","az")
```

## Part 1

Over time, the one that stays the closest is among the ones that has the lowest acceleration,
if more than one, then the one with the lowest velocity,
if more than one, then the one with the lowest position

```{r,eval=FALSE}
### find the manhattan acceleration
particleframe<-particleframe %>% rowwise %>%
  ##find the manhattan accelerations
  mutate(mana=abs(ax)+abs(ay)+abs(az))%>%
  ##find starting manhattan manhattan velocity
  mutate(manv=abs(vx)+abs(vy)+abs(vz))%>%
  ##find starting manhattan distance
  mutate(manp=abs(px)+abs(py)+abs(pz))
```


```{r,eval=FALSE}

part1<-particleframe%>%rowwise%>%
  filter(mana==min(particleframe$mana))
part1<-part1%>%rowwise%>%
  filter(manv==min(part1$manv))
part1<-part1%>%rowwise%>%
  filter(manp==min(part1$manp))
part1<-part1$num[1]
part1
```
## Part 2

After spending a lot of time trying to find and calculate the collisions  AND GETTING WRONG ANSWERS, just step through.
For time from 1-50 (the calculations say they end at 39, this is just to be certain), calculate position.  Remove any particles at the same position.

```{r,eval=FALSE}
killcolliders<-function(df){
  dead<-list()
  ## copy the ps and vs
  df<-df %>% rowwise %>%mutate(base1=num+1)%>%
    ## add working copies of position and velocity
    mutate(curvx=vx,curvy=vy,curvz=vz,curpx=px,curpy=py,curpz=pz)
  i<-1
  while(i<50){
  df<-df%>%rowwise%>%
    ## update velocity
    mutate(nxtvx=curvx+ax,nxtvy=curvy+ay,nxtvz=curvz+az)%>%
    ## update position
    mutate(nxtpx=nxtvx+curpx,nxtpy=nxtvy+curpy,nxtpz=nxtvz+curpz)%>%
    ## create an easy-to-compare coordinate
    mutate(cmpr=paste(nxtpx,nxtpy,nxtpz))
  ## get a list of everything that appears more than once
  collided<-df$cmpr[which(duplicated(df$cmpr))]
  ## mark everything that has collided
  df<-df%>%rowwise%>%
    mutate(clsn=cmpr %in% collided)
  ## dead is for seeing the collisions later on, if wanted
  #dead<-append(dead,list(df%>%rowwise%>%filter(clsn==TRUE)))
  ## filter out the things that haven't banged into anything
  df<-df%>%rowwise%>%
    filter(clsn==FALSE)%>%
    ### update position, and velocity
    mutate(curvx=nxtvx,curvy=nxtvy,curvz=nxtvz,curpx=nxtpx,curpy=nxtpy,curpz=nxtpz)
  ## update the counter
  i<-i+1}
  df}
```




```{r,eval=FALSE}
part2<-killcolliders(particleframe)
part2<-nrow(part2)
part2
```

---------
Everything below here works, but takes actual hours.  HOURS. I still think this is the right way to do it (find the actual collisions), but HOURS.  (Also, it didn't work at first, which is why it was abandoned.)

```{r,eval=FALSE}
# collisiontime<-function(df){
#   ## find t where the xs are equal
#   ## because of the way this moves, the formula isn't technically
#   ## a*t^2+v*t+p, it's (a/2)*t^2+(v+a/2)*t+p
#   a<-(df$ax[1]/2)-(df$ax[2]/2)
#   b<-(df$vx[1]+df$ax[1]/2)-(df$vx[2]+df$ax[2]/2)
#   c<-df$px[1]-df$px[2]
### if a=0, then use the formula for a line
#   if(a==0){
### if a & b=0, then these will never meet, return NA
#     if(b==0){return(NA)}else{t<-(-c/b)}
#   }else{
### find the determinant - if less than 0, no intersection
#     d<-b^2-(4*a*c)
#     if(d<0){t<-return(NA)
#     }else{
### ts are the 0s
#       t<-c(((-b-sqrt(d))/(2*a)),((-b+sqrt(d))/(2*a)))}}
### filter if t<0 or t is not an integer
#   t<-t[which(t>=0)]
#   t<-t[which(t==round(t))]
#   if(length(t)==0){return(NA)}
### ensure that it intersects in the y & z coordinates as well
#   t<-sapply(t,function(x){
#     if(((df$ay[1]/2)*x^2+(df$ay[1]/2+df$vy[1])*x+df$py[1])==((df$ay[2]/2)*x^2+(df$ay[2]/2+df$vy[2])*x+df$py[2])){x<-x}else{return(NA)}
#     if(((df$az[1]/2)*x^2+(df$az[1]/2+df$vz[1])*x+df$pz[1])==((df$az[2]/2)*x^2+(df$az[2]/2+df$vz[2])*x+df$pz[2])){x<-x}else{return(NA)}
#     x})
#   t<-t[which(!is.na(t))]
#   if(length(t)==0){return(NA)}
### return time t (or ts if multiple intersections)
#   t}

```


```{r,eval=FALSE}
# collisionframeagain<-as.data.frame(matrix(nrow=0,ncol=3))
# for(i in 1:(nrow(particleframe)-1)){
#   for(j in (i+1):nrow(particleframe)){
### for each pair of particles, find out if they intersect or not
#     x<-collisiontime(particleframe[c(i,j),])
### create a data frame with their intersection time (if it exists) or NA (if not)
#     for(k in 1:length(x)){
#       collisionframeagain<-rbind(collisionframeagain,(c(i,j,x[k])))}}
#   if(i%%100==0){cat(i,"\n")}}
```




```{r,eval=FALSE}
#colnames(collisionframeagain)<-c("A","B","Collide")
### pull the list of uncollided particles and count them.
#answer<-length(unique(c(collisionframeagain$A,finalshot$B)))
```

```{r,eval=FALSE}
### get just the collisions
#justthecollisions<-collisionframeagain%>%rowwise%>%filter(!is.na(Collide))
```


```{r,eval=FALSE}
#  remcol<-function(df){
# ### get the list of particles that collided
#  hitlist<-c()
# ### if there are still particles that have a collision time
#    while(any(!is.na(df$Collide))){
# ###   get a sub df of all with the minimum collision time
#      b<-df%>%rowwise%>%filter(Collide==min(df$Collide,na.rm=TRUE))
# ### get a list of all particles that collided 
#      c<-unique(c(b$A,b$B))
# ### 
# hitlist<-c(hitlist,c)
# ### remove those particles from further consideration, repeat
#      df<-df%>%rowwise%>%filter(!(A%in%c))%>%filter(!(B%in%c))}
#    hitlist}
```



```{r}
#hitparticles<-remcol(justthecollisions)
### answer is the number of particles - the length of the particles that collided.
#answer<-nrow(particleframe)-length(hitparticles)
#answer
```


