---
title: "Day 11 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}

library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readr)
library(collections)
options(scipen = 999)
```

```{r}
input<-read_lines("Day11Sample.txt")
```

## Part 1
I blatantly stole this one:

Using three coordinates (instead of two) it's possible to figure out how far away - 
for this:
a is N/S. Moving N/S a doesn't change.  Moving E, a goes down, moving W, a goes up
b is NE/SW.  Moving SE or S, b goes up, moving N or NW, b goes down
c is NW/SE.  Moving N/NE, c goes up, moving S/Sw, c goes down.
so, every move, one thing is going up, one is going down.
the total distance, will be the max of the absolute value of a, b, and c

```{r}
hexed<-function(str){
  str<-unlist(str_split(str,","))
  a<-0
  b<-0
  c<-0
  for(i in 1:length(str)){
    stp<-str[i]
    switch(stp,
           "n"={b<-b-1
           c<-c+1},
           "ne"={a<-a-1
           c<-c+1},
           "se"={a<-a-1
           b<-b+1},
           "s"={b<-b+1
           c<-c-1},
           "sw"={a<-a+1
           c<-c-1},
           "nw"={a<-a+1
           b<-b-1},
           cat(stp,"-something is wrong\n"))}
  answer<-max(abs(a),abs(b),abs(c))
  answer}

```


```{r}
sapply(input,hexed)

```
## Part 2
Small change to save the max as it goes by:
```{r}
veryhexed<-function(str){
  str<-unlist(str_split(str,","))
  a<-0
  b<-0
  c<-0
  maxabc<-0
  for(i in 1:length(str)){
    stp<-str[i]
    switch(stp,
           "n"={b<-b-1
           c<-c+1},
           "ne"={a<-a-1
           c<-c+1},
           "se"={a<-a-1
           b<-b+1},
           "s"={b<-b+1
           c<-c-1},
           "sw"={a<-a+1
           c<-c-1},
           "nw"={a<-a+1
           b<-b-1},
           cat(stp,"-something is wrong\n"))
    maxabc<-max(maxabc,max(abs(a)),max(abs(b)),max(abs(c)))}
  answer<-max(abs(a),abs(b),abs(c))
  maxabc}

```

```{r}
sapply(input,veryhexed)

```

