---
title: "Day 10 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringi)
library(stringr)
library(tidyverse)
library(readr)
library(collections)
options(scipen = 999)
```

```{r}
input<-c(3, 4, 1, 5)
twisthere<-as.numeric(unlist(str_split(input,",")))
```

## Part 1
So, need a function that takes in the list of lengths & the list of numbers

```{r}
twistme<-function(twists=input,circle=0:4,reps){
  ### set the skip length to 0
  skp<-0
  now<-1
  ### repeats added for part 2
  for(j in 1:reps){
  for(i in 1:length(twists)){
    tw<-twists[i]
    ### if it doesn't need to wrap around
    if(tw+now-1<=length(circle)){
      circle[now:(now+tw-1)]<-rev(circle[now:(now+tw-1)])
    }else{
      ### get the vector to reverse (the end of this one, the first part of the next)
      sw<-c(circle[now:length(circle)],circle[1:(now+tw-1-length(circle))])
      ### reverse it
      sw<-rev(sw)
      ### paste it in
      circle[now:length(circle)]<-sw[1:(length(circle)-now+1)]
      circle[1:(now+tw-1-length(circle))]<-sw[(length(circle)-now+2):length(sw)]}
    now<-now+tw+skp
    while(now>length(circle)){now<-now-length(circle)}
    skp<-skp+1}}
circle}
```

```{r}
p1<-twistme(twisthere,0:255,1)
part1<-prod(p1[1:2])
part1

```

##Part 2

First - small change in the "twistme()" function to do multiple repeats
Second - convert the input string


```{r}
### this is just to show what conversion is supposed to do
twisttwo<-c(sapply(unlist(str_split(input,"")),utf8ToInt),c(17,31,73,47,23))
```
Third - create a very specific xor that works on 16 decimals to change them into another decimal.

```{r}
xordec<-function(n,m){
  answer<-0
  for(i in 15:0){
    ### converts left to right by figuring out if it has that power of 2, then moving down.
    answer<-answer+xor((n%/%(2^i)),(m%/%(2^i)))*2^i
    n<-n%%2^i
    m<-m%%2^i}
  answer}
```

Next - create a function that takes a sparse hash and makes it dense

```{r}
### This ONLY works well if there are a multiple 16 characters. ONLY
densehash<-function(sphsh){
  dh<-c()
  while(length(sphsh)>0){
    clump<-sphsh[1]
    ## start xoring the first numbers from 1-15
    for(i in 2:16){
      clump<-xordec(clump,sphsh[i])}
    ## remove those numbers from the list
    sphsh<-sphsh[-(1:16)]
    ## add that to the outpu
    dh<-c(dh,clump)}
dh}
```

penultimately - convert a dense hash into a knot hash

```{r}
knothash<-function(dh){
  kh<-(sapply(dh,function(x){
    ## change each number to hexmode and then to a character
    x<-as.character(as.hexmode(x))
    ## if it is <=15, make sure that there are two digits - add a 0 at the front
    if(nchar(x)==1){x<-str_c("0",x)}else{x}
    x}))
  ##smush everything together
  kh<-str_flatten(kh)
kh}
```

Then to run it all at once:

```{r}
### Remember, the input is a string
input<-("1,2,3")
reviselength<-unlist(c(sapply(unlist(str_split(input,"")),utf8ToInt),c(17,31,73,47,23)))
sparsehash<-twistme(reviselength,0:255,64)
dh<-densehash(sparsehash)
part2<-knothash(dh)
part2
```